---
title: "Reference file parse trial"
output: html_notebook
---

Experiments for parsing reference files.

Load the tidyverse and set the reference file name.

```{r}
library(tidyverse, quietly = TRUE)
ref <- "combined.ref"
```

The reference file consists of two sections: The information about the code
compiled and the information about the symbols in the GAMS program
(aka symbol table). The information about the code compiled consists of
records with 11 fields. The meaning of the fields is:

- record count
- symbol index
- symbol name
- symbol type
- reference type (available: declared, defined, impl-asn, assigned, ref, control)
- global listing line
- local line
- local column position (due to macros this many not be always 100% correct)
- include nesting level
- index in the file summary
- file name

```{r}
code <- read_delim(ref,
                   delim = " ",
                   col_names = c("record_count",
                                 "symbol_index",
                                 "symbol_name",
                                 "symbol_type",
                                 "reference_type",
                                 "global_listing_line",
                                 "local_line",
                                 "local_column_position",
                                 "include_nesting_level",
                                 "index_in_the_file_summary",
                                 "file_name")
             ) %>%
filter(symbol_type != 'FUNCT') %>%
filter(symbol_type %in% c('SETS', 'PARAMETERS'))
View(code)
```

The symbol table starts with the line that is indicated by a 0 in the first
field, followed by the number of records in this symbol table and the text
"size of symboltable".

```{r}
header <- read_delim(ref,
                skip = nrow(code),
                n_max = 1,
                delim = " ",
                col_names = c("zero",
                              "symbols",
                              "description")
               ) %>%
filter(zero == '0')
View(header)
```

The records for the symbol table consists of:

- internal symbol index
- symbol name
- numerical and string symbol type
- dimension (0 for functions and models)
- cardinality (0 for functions, card(sym) for regular symbols, and number of equation symbols for models) 
- a list of length dimension with the internal symbol index of the domain set (0 if universe)
- the symbol text

```{r}
symbol <- read_delim(ref,
                     skip = nrow(code)+nrow(header),
                     delim = " ",
                     col_names = c("internal_symbol_index",
                                   "symbol_name",
                                   "symbol_type_code",
                                   "symbol_type",
                                   "dim",
                                   "cardinality",
                                   "dim1",
                                   "dim2",
                                   "dim3",
                                   "dim4",
                                   "dim5",
                                   "dim6",
                                   "dim7",
                                   "dim8",
                                   "dim9"
                                  )
                    ) %>%
filter(symbol_type != 'FUNCT') %>%
filter(symbol_type != 'PRED')
View(symbol)
```
Switch to Python and import some useful packages.
```{python}
import pandas
from pprint import pprint
import re
import sys
print(f"The python version is {sys.version}")
```
Define regular expression based fields-from-line parsing function
```{python}
def parse(line, fields):
  # Use the first field dictionary to cache the compiled regular expression
  if 'regex' not in fields[0]:
    # Construct the regular expression from the field captures once only
    fields[0]['regex'] = re.compile("(" + ")\\s+(".join([f['capture'] for f in fields]) + ")$")
  ma = fields[0]['regex'].match(line)
  if (ma is None):
    return None
  # Store the matched fields in a record dictionary
  record = {r['name']:ma[i+1] for i,r in enumerate(fields)}
  # Convert integer fields in record to int
  for f in fields:
    if f['capture'] == '\\d+':
      record[f['name']] = int(record[f['name']])
  return record
```

Define categories and capturing regular expressions for parsing.
```{python}
code_fields = [
  {'name': 'record_count',              'capture': '\\d+'},
  {'name': 'symbol_index',              'capture': '\\d+'},
  {'name': 'symbol_name',               'capture': '[^\\s]+'},
  {'name': 'symbol_type',               'capture': '[^\\s]+'},
  {'name': 'reference_type',            'capture': '[^\\s]+'},
  {'name': 'global_listing_line',       'capture': '\\d+'},
  {'name': 'local_line',                'capture': '\\d+'},
  {'name': 'local_column_position',     'capture': '\\d+'},
  {'name': 'include_nesting_level',     'capture': '\\d+'},
  {'name': 'index_in_the_file_summary', 'capture': '\\d+'},
  {'name': 'file_name',                 'capture': '.*'}
]
header_fields = [
  {'name': 'zero',        'capture': '\\d+'},
  {'name': 'symbols',     'capture': '\\d+'},
  {'name': 'description', 'capture': 'size of symboltable'}
]
symbol_fields = [
  {'name': 'internal_index', 'capture': '\\d+'},
  {'name': 'name',           'capture': '[^\\s]+'},
  {'name': 'type_code',      'capture': '\\d+'},
  {'name': 'type',           'capture': '[^\\s]+'},
  {'name': 'dim',            'capture': '\\d+'},
  {'name': 'cardinality',    'capture': '\\d+'},
  {'name': 'remainder',      'capture': '.*'}
]
```
Parse the reference fie, code references first, then the header and body of the symbol table
```{python}
#filepath = 'combined.ref'
filepath = '../../Trunk_build_artifacts/Model/4_model.ref'
with open(filepath) as fp:
  # Parse the code references
  code_refs = []
  while True:
    # Read code reference record line and match fields
    line = fp.readline()
    code = parse(line, code_fields)
    if (code is None):
      break
    code_refs.append(code)
    assert code['record_count'] == len(code_refs) # verify that record count is sequential and hence implied by list index
    del code['record_count'] # drop superfluous count
    assert code['symbol_index'] > 0 # verify that the symbol index makes sense
    code['symbol_index'] -= 1 # convert 1-based index to 0-based
  print(f"Parsed {len(code_refs)} code reference records")

  # Parse the symbol table header in the same manner
  header = parse(line, header_fields)
  if (header is None):
    raise RuntimeError(f"Could not match header field! Regexp {header_regex} did not match line '{line}'")
  print(f"Parsed symbol table header: {header}")

  # Parse the symbol table
  symbols = []
  while True:
    # Read symbol records line and match fields
    line = fp.readline()
    symbol = parse(line, symbol_fields)
    if (symbol is None):
      break
    symbols.append(symbol)
    assert symbol['internal_index'] == len(symbols) # verify that list indexing (0-based) can be used instead
  assert len(symbols) == header['symbols']
  print(f"Parsed {len(symbols)} symbol records")
```
The remaining part of the symbol records has a variable number of fields, parse these bits separately.
```{python}
remainder_regexes = {}
for s in symbols:
  if s['dim'] == 0:
    s['description'] = s['remainder']
  else:
    if s['dim'] not in remainder_regexes:
      remainder_regexes[s['dim']] = re.compile("\\s+".join(["(\\d+)"]*s['dim']) + "\\s+(.*)$")
    ma = remainder_regexes[s['dim']].match(s['remainder'])
    assert ma is not None
    ma_list = [ma[i] for i in range(1, s['dim']+2)]
    s['domain_sets'] = [int(m)-1 for m in ma_list[:-1]] # convert 1-based index to 0-based, with -1 == *
    s['description'] = ma_list[-1]
```
Remove superfluous fields from the symbol dictionaries
```{python}
for s in symbols:
  del s['remainder'] # parsed to domain_sets and description
  del s['dim'] # implied by presence and length of domain_sets
  if s['description'] == "":
    del s['description'] # description absent, drop field
  del s['internal_index'] # implied by symbols list index
```
Verify that domain set indexes reference sets
```{python}
for s in symbols:
  if 'domain_sets' in s:
    for d in s['domain_sets']:
      if d >= 0:
        assert symbols[d]['type'] == 'SET'
```
Show unique type values
```{python}
print("Code references 'reference_type':", set([c['reference_type'] for c in code_refs]))
print("Code references 'symbol_type':", set([c['symbol_type'] for c in code_refs]))
print("Symbols 'type':", set([s['type'] for s in symbols]))
```
Display the code references of a particular type
```{python}
shown = 0
for i,c in enumerate(code_refs):
  if c['reference_type'] == 'assigned':
    print(i)
    pprint(c)
    shown += 1
    if shown >= 30:
      print("Show limit reached!")
      break
```
Display the symbol table
```{python}
shown = 0
for i,s in enumerate(symbols):
  if s['type'] == 'PARAM':
    print(i)
    pprint(s)
    shown += 1
    if shown >= 40:
      print("Show limit reached!")
      break
```

